\LARGE Untangling the link between traits, size and growth rate in plants

### 2016-07-28 ms/ms.tex
Standardise maths with Falster et al 2016, Gibert et al 2016

Eqs. \ref{eq:dhdt}-\ref{eq:dDdt} are mathematically true, and must
therefore hold for any model of plant growth. To make explicit
predictions, however, requires two additional elements: 1) A specific
model describing how the various mass and area terms
($m_\textrm{l}, m_\textrm{s}, m_\textrm{b}, m_\textrm{r}, a_\textrm{l}, a_\textrm{s}, a_\textrm{b}$)
vary relative to one another; and 2) For any trade-offs in plant
function to be specified; this is where traits have an influence.  These two components are outlined in following sections.

@@

We begin with a standard model for the amount of biomass available for
growth, $\textrm{d}b / \textrm{d}t$, given by the difference between income
(total photosynthesis) and losses (respiration and turnover) within the
plant \citep{makela-1997, Thornley-2000, falster-2011}:

\begin{equation}\label{eq:dbdt}
\underbrace{\strut \frac{\textrm{d}b}{\textrm{d}t}}_\textrm{net biomass production}
  = \underbrace{\strut y}_\textrm{yield}
    \big( \underbrace{\strut a_\textrm{l} \, p}_\textrm{photosynthesis} -
     \underbrace{\strut \,\sum_\textrm{i=l,b,s,r}{m_\textrm{i} \, r_\textrm{i}}}_\textrm{respiration}\big)
    - \underbrace{\strut \sum_\textrm{i=l,b,s,r}{m_\textrm{i} \, k_\textrm{i}}}_\textrm{turnover}.
\end{equation}

Here, $m,r$, and $k$ refer to the mass, respiration rate, and
turnover rate of different tissues, denoted by subscripts $l$=leaves,
$b$=bark, $s$=sapwood and $r$=roots. $A$ is the assimilation
rate of CO$_2$ per leaf area and $y$ is yield: the fraction of
assimilated carbon fixed in biomass (the remaining fraction being lost
as growth respiration) (see Table \ref{tab:params} for units and
definitions). Photosynthesis is proportional to leaf area,
$a_\textrm{l} = m_\textrm{l} / \phi$, where $\phi$ is leaf mass per area.
The total mass of living tissues $m_\textrm{a}=m_\textrm{l}+m_\textrm{b}+m_\textrm{s}+m_\textrm{r}.$

@@

\begin{equation} \label{eq:dhdt}
\underbrace{\strut\frac{\textrm{d}h}{\textrm{d}t}}_\textrm{height growth rate}= \underbrace{\strut\frac{\textrm{d}h}{\textrm{d}a_\textrm{l}}}_\textrm{architecture}
\times \underbrace{\strut\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{a}}}_\textrm{marginal leaf deployment}
\times \underbrace{\strut\frac{\textrm{d}m_\textrm{a}}{\textrm{d}b}}_\textrm{allocation to growth}
\times \underbrace{\strut\frac{\textrm{d}b}{\textrm{d}t}}_\textrm{mass production}.
\end{equation}

The first term on the right of eq \ref{eq:dhdt},
$\textrm{d}h / \textrm{d}a_\textrm{l}$, is the growth in plant height
per unit growth in total leaf area -- accounting for the architectural
strategy of the plant. Some species tend to leaf out more than grow
tall, while other species emphasise vertical
extension \citep{poorter-2006}.

The second term, $\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{a}$,
accounts for the marginal cost of deploying an additional unit of leaf
area, including construction of the leaf itself and various support
structures. As such, $\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{a}$
can itself be expressed as a sum of construction costs per unit leaf
area:

\begin{equation}\label{eq:daldmt}
\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{a}}
= \big(\phi
 + \frac{\textrm{d}m_\textrm{s}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}m_\textrm{b}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}}\big)^{-1}.
\end{equation}

The third term in eq \ref{eq:dhdt},
$\textrm{d}m_\textrm{a} / \textrm{d}P$, gives the fraction of net
biomass production (eq. \ref{eq:dbdt}) that is allocated to growth
rather than reproduction or storage.

In a similar way, basal area ($a_\textrm{st}$) increment can be
expressed as the sum of increments in sapwood, bark \& heartwood areas
($a_\textrm{ss}, a_\textrm{sb}, a_\textrm{sh}$ respectively):
$\frac{\textrm{d}a_\textrm{st}}{\textrm{d}t}= \frac{\textrm{d}a_\textrm{sb}}{\textrm{d}t} + \frac{\textrm{d}a_\textrm{ss}}{\textrm{d}t} + \frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t}$.
We now have an equation for basal area growth that conatins many of the same
elements as eq. \ref{eq:dhdt}:

\begin{equation}\label{eq:dast}
\frac{\textrm{d}a_\textrm{st}}{\textrm{d}t}=
\left(\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}a_\textrm{l}} + \frac{\textrm{d}a_\textrm{sb}}{\textrm{d}a_\textrm{l}}\right) \times
\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{a}} \times \frac{\textrm{d}m_\textrm{a}}{\textrm{d}b} \times \frac{\textrm{d}b}{\textrm{d}t} +\frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t} .
\end{equation}

Diameter growth is then given by the geometric relationship between stem
diameter ($D$) and $a_\textrm{st}$:

\begin{equation} \label{eq:dDdt}
\frac{\textrm{d}D}{\textrm{d}t}= \sqrt{\frac{\pi}{a_\textrm{st}}} \frac{\textrm{d}a_\textrm{st}}{\textrm{d}t}.
\end{equation}

@@

By definition, the main effect of increasing $\phi$ is to decrease the
marginal leaf deployment per mass (eq. \ref{eq:daldmt}) and thus reduce
growth rate. However, to fully account for the influence of $\phi$ on
growth we must also account for the decrease in leaf turnover that
results from superior leaf construction \citep{wright-2004}.
This is achieved by linking the turnover rate of leaf in eq.
\ref{eq:dbdt} to $\phi$. We use an allometric scaling relationship of
the form $k_\textrm{l}=\alpha_4 \, \phi^{-\beta_4}$, which has been
shown to describe patterns across diverse
ecosystems \citep{wright-2004} (Fig. \ref{fS-leaf}).

@@

Unlike leaf and stem traits, height at maturation ($h_m$) moderates
growth by adjusting the amount of energy invested in growth, i.e.~the
term $\frac{\textrm{d}m_\textrm{a}}{\textrm{d}b}$ in eqs.
\ref{eq:dhdt} and \ref{eq:dast}.

@@

Unlike eq. \ref{eq:RGR}, eqs. \ref{eq:dhdt}, \ref{eq:dast} and
\ref{eq:dDdt} predict a relationship between $\phi$ and growth that
changes with plant size ( \ref{f-lma_growth_size}).
Decreasing $\phi$ has two impacts on growth rate. First, lower
$\phi$ increases marginal leaf deployment
($\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{a}$ by economising on
construction costs. Second, lower $\phi$ decreases net production
($\textrm{d}P / \textrm{d}t$), due to increased leaf turnover. Whether
lower $\phi$ increases growth thus depends on the relative magnitude
of these two effects. When plants are small the effect on leaf
deployment rate is bigger and so decreasing $\phi$ increases growth
rate. When plants are large, the influence of $\phi$ on leaf
deployment rate is diminished, because the costs of building other
supportive tissues (other terms in eq. \ref{eq:daldmt}) are larger (Fig.
\ref{f-mass_fraction}). The net result is that at larger sizes, low
$\phi$ is no longer advantageous for growth (Fig.
\ref{f-lma_growth_size}d).

@@

 Greater height at maturation ($h_m$)
can thus lead to a growth advantage by increasing
$\frac{\textrm{d}m_\textrm{a}}{\textrm{d}b}$ . At smaller sizes, there is no differentiation
among species, because most species are focusing on growth. At larger sizes, individuals of some species are
allocating a majority of their surplus energy to growth, leaving a
strong signal in potential growth rate.

@@

Current theory emphasises the relationship between $\phi$ and the relative growth in plant mass
\citep{lambers-1992, cornelissen-1996, wright_cross-2000,  enquist-2007},
$\textrm{d}P / \textrm{d}t / m_\textrm{a}$. This relationship can be derived from eq. \ref{eq:dbdt}  by ...
(see Appendix \ref{app:traits-RGR}).

@@

\begin{table}[ht]
 \caption{Key variables
%  $a = l + ss + sb +  r$,
%  $m_{sh}$,
%  $m_t = m_a + m_{sh}$,
%  $st = stem total$
}
\centering

@@

  $b$   & kg  & biomass originating from parent plant\\
  $m_i$ & kg  & mass retained on plant in tissue $i$\\
  $a_i$ & m$^2$  & surface-area or cross-sectional-area of tissue $i$\\
  $h$   & m  & height of plant\\
  $D$   & m  & diameter of plant stem\\

@@

{\footnotesize  % smaller font, double space
 \begin{doublespace}
  \begin{tabular}{p{2.5cm}p{3.5cm}p{5cm}p{4cm} }\\ \hline
  Variable & Function & Marginal cost & Growth rate\\ \hline
  \multicolumn{4}{l}{\textbf{a) Assumed relationship to leaf area}} \\ \\
  height &
    $h = \alpha_1 \, a_\textrm{l}^{\beta_1}$ &
    $\frac{\textrm{d}h}{\textrm{d}a_\textrm{l}} =  \beta_1 \alpha_1 a_\textrm{l}^{\beta_1-1}$ &
    $\frac{\textrm{d}h}{\textrm{d}t}  = \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\
  sapwoood area &
    $a_\textrm{ss} = \theta \, a_\textrm{l}$ &
    $\frac{\textrm{d}a_\textrm{ss}}{\textrm{d} a_\textrm{l}} = \theta$ &
    $\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}t}  =\frac{\textrm{d}a_\textrm{ss}}{\textrm{d} a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\
  bark area &
    $a_\textrm{sb} = b \, \theta \, a_\textrm{l}$ &
    $\frac{\textrm{d}a_\textrm{sb}}{\textrm{d} a_\textrm{l}} = b \, \theta$ &
    $\frac{\textrm{d}a_\textrm{sb}}{\textrm{d}t} = \frac{\textrm{d}a_\textrm{sb}}{\textrm{d} a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\  \\
  \multicolumn{4}{l}{\textbf{b) Derived equation for mass of tissue }} \\
  leaf mass &
    $m_\textrm{l} = \phi \, a_\textrm{l} $ &
    $\frac{\textrm{d}m_\textrm{l}}{\textrm{d}a_\textrm{l}} = \phi$ &
    $\frac{\textrm{d}m_\textrm{l}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{l}}{\textrm{d}a_\textrm{l}}  \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\
  sapwood mass &
    $m_\textrm{ss} = \theta \, \rho \, \eta_c \, a_\textrm{l} \, h $ &
    $\frac{\textrm{d}m_\textrm{ss}}{\textrm{d}a_\textrm{l}} = \theta\, \rho\, \eta_c\, \big( h + a_\textrm{l}\, \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}} \big)$ &
    $\frac{\textrm{d}m_\textrm{ss}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{ss}}{\textrm{d}a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\
  bark mass &
    $m_\textrm{sb} = b\, \theta \, \rho \, \eta_c \, a_\textrm{l} \, h $ &
    $\frac{\textrm{d}m_\textrm{sb}}{\textrm{d}a_\textrm{l}} = b \, \theta \, \rho \, \eta_c\big( h + a_\textrm{l} \, \frac{\textrm{d}h}{\textrm{d}a_\textrm{l}} \big)$ &
    $\frac{\textrm{d}m_\textrm{sb}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{sb}}{\textrm{d}a_\textrm{l}} \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\
  root mass &
    $m_\textrm{r} = \alpha_3 \, a_\textrm{l}$ &
    $\frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}} = \alpha_3$  &
    $\frac{\textrm{d}m_\textrm{r}}{\textrm{d}t}  = \frac{\textrm{d}m_\textrm{r}}{\textrm{d}a_\textrm{l}}  \, \frac{\textrm{d}a_\textrm{l}}{\textrm{d}t}$ \\ \\

  \multicolumn{4}{l}{\textbf{c) Integrated quantities}} \\
  heartwood area &
    $a_\textrm{sh} = \int_0^t \frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t}(t^\prime) \, dt^\prime$ &
     &
    $\frac{\textrm{d}a_\textrm{sh}}{\textrm{d}t} = k_\textrm{s} \, a_\textrm{ss}$ \\
  heartwood mass &
    $m_\textrm{sh} = \int_0^t \frac{\textrm{d}m_\textrm{sh}}{\textrm{d}t}(t^\prime) \, dt^\prime$ &
     &
    $\frac{\textrm{d}m_\textrm{sh}}{\textrm{d}t} = k_\textrm{s} \, m_\textrm{ss}$ \\
  \hline \\
\end{tabular}
NB -  theta is sapwood area per leaf area. In tree 2 has been opposite, but that makes notation more clumsy.
\end{doublespace}}
\label{tab:allometry}
\end{table}

@@

\begin{table}[ht]
\caption{Effect of traits and related trade-offs on demography}
\centering
{\footnotesize  % smaller font, double space
  \begin{doublespace}
  \begin{tabular}[c]{l|c|ccccc|c|c}
  \multicolumn{9}{c}{}\\ \hline
  & \multicolumn{8}{c}{\textbf{Effect of trait on}}\\
  & \textbf{Birth} & \multicolumn{5}{c|}{\textbf{Growth}} & \textbf{Mortality} & \textbf{Fecundity} \\
  & $h_0$
  & $\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{a}}$
  & $\frac{\textrm{d}m_\textrm{a}}{\textrm{d}b}$
  & $\frac{\textrm{d}b}{\textrm{d}t}$
  & $\frac{\textrm{d}a_\textrm{ss}}{\textrm{d}a_\textrm{l}}$
  & $\frac{1}{k_s}$ & &  \\\hline
  \textbf{This study}&&&&&&&&\\
  Leaf-construction cost, $\phi$ & & $\downarrow$ & $\uparrow$ & & & & \\
  Wood-construction cost, $\rho$ & & $\downarrow$ &  & & & $\downarrow$ & $\downarrow$  \\
  Height at maturation, $h_m$ & & &$\uparrow$ & & & & & $\downarrow$  \\
  &&&&&&&&\\\hline
  \textbf{Possible extensions}&&&&&&&&\\
  Sapwood area per leaf area, $\theta$ & & $\downarrow$& & $\uparrow$ & $\uparrow$ & & \\
  Leaf-nitrogen per area & & &$\downarrow$$\uparrow$ & & & & & \\
  Seed size & $\uparrow$ & & & & & & & $\downarrow$ \\ \hline
  \end{tabular}
  \end{doublespace}
}
\label{tab:trade-offs}
\end{table}

@@

We start with eq. \ref{eq:dbdt} for mass-based growth. For seedlings, which are
young and mostly leaf, it is reasonable to ignore all turnover terms as
well as the respiration terms for non-leaf tissues. Net production then
becomes a linear function of leaf area and net photosynthesis per leaf
area ($A_\textrm{net} = y(A - r_\textrm{l})$), making relative growth
rate a linear function of $\phi$:

\begin{equation}\label{eq:RGR}
\underbrace{\strut\frac{\textrm{d}b}{\textrm{d}t}\frac{1}{m_\textrm{a}}}_\textrm{relative growth in mass}  \approx A_\textrm{net} \times \phi^{-1} \times \frac{m_\textrm{l}}{m_\textrm{a}}. \end{equation}

Although Eq. \ref{eq:RGR} captures patterns of growth in seedlings in
relation to $\phi$\citep{wright_cross-2000}, this
approximation does not directly link to other traits, or to the
variables that are routinely collected for large trees: namely plant
height ($h$) and stem cross-sectional area ($a_\textrm{st}$) or
diameter $D$.

@@

\begin{equation}\label{eq:ha}
a_\textrm{l}=\alpha_1 \, h^{\beta_1}.
\end{equation}

@@

\begin{equation}\label{eq:crown2}
\frac{m_\textrm{l}}{\phi}= \theta \, S(0,h).
\end{equation}

@@

\begin{equation}\label{eq:ms1}
m_\textrm{s}=\rho \, \int_0^h \, S(z,h) \, \textrm{d}z= \rho \, S(0,h) \, h \, \eta_c, \end{equation}

@@

\begin{equation}\label{eq:ms2}
m_\textrm{s}=\rho \, \eta_c \, \theta \, a_\textrm{l} \, h.
\end{equation}

@@

\begin{equation}\label{eq:mb}
m_\textrm{b}=b m_\textrm{s}.
\end{equation}

@@

\begin{equation}\label{eq:mr}
m_\textrm{r}=\alpha_3 \, a_\textrm{l}.
\end{equation}

>>>>

Eqs. \ref{eq:dhdt}-\ref{eq:dDdt} are mathematically true, and must
therefore hold for any model of plant growth. To make explicit
predictions, however, requires two additional elements: 1) A specific
model describing how the various mass and area terms
($M_{\rm l}, M_\textrm{s}, M_{\rm b}, M_{\rm r}, A_{\rm l}, A_\textrm{s}, A_{\rm b}$)
vary relative to one another; and 2) For any trade-offs in plant
function to be specified; this is where traits have an influence.  These two components are outlined in following sections.

@@

We begin with a standard model for the amount of biomass available for
growth, ${\rm d}B / {\rm d}t$, given by the difference between income
(total photosynthesis) and losses (respiration and turnover) within the
plant \citep{makela-1997, Thornley-2000, falster-2011}:

\begin{equation}\label{eq:dbdt}
\underbrace{\strut \frac{{\rm d}B}{{\rm d}t}}_\textrm{net biomass production}
  = \underbrace{\strut y}_\textrm{yield}
    \big( \underbrace{\strut A_{\rm l} \, p}_\textrm{photosynthesis} -
     \underbrace{\strut \,\sum_\textrm{i=l,b,s,r}{M_\textrm{i} \, r_\textrm{i}}}_\textrm{respiration}\big)
    - \underbrace{\strut \sum_\textrm{i=l,b,s,r}{M_\textrm{i} \, k_\textrm{i}}}_\textrm{turnover}.
\end{equation}

Here, $m,r$, and $k$ refer to the mass, respiration rate, and
turnover rate of different tissues, denoted by subscripts $l$=leaves,
$b$=bark, $s$=sapwood and $r$=roots. $A$ is the assimilation
rate of CO$_2$ per leaf area and $y$ is yield: the fraction of
assimilated carbon fixed in biomass (the remaining fraction being lost
as growth respiration) (see Table \ref{tab:params} for units and
definitions). Photosynthesis is proportional to leaf area,
$A_{\rm l} = M_{\rm l} / \phi$, where $\phi$ is leaf mass per area.
The total mass of living tissues $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_\textrm{s}+M_{\rm r}.$

@@

\begin{equation} \label{eq:dhdt}
\underbrace{\strut\frac{{\rm d}H}{{\rm d}t}}_\textrm{height growth rate}= \underbrace{\strut\frac{{\rm d}H}{{\rm d}A_{\rm l}}}_\textrm{architecture}
\times \underbrace{\strut\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}}_\textrm{marginal leaf deployment}
\times \underbrace{\strut\frac{{\rm d}M_{\rm a}}{{\rm d}B}}_\textrm{allocation to growth}
\times \underbrace{\strut\frac{{\rm d}B}{{\rm d}t}}_\textrm{mass production}.
\end{equation}

The first term on the right of eq \ref{eq:dhdt},
${\rm d}H / {\rm d}A_{\rm l}$, is the growth in plant height
per unit growth in total leaf area -- accounting for the architectural
strategy of the plant. Some species tend to leaf out more than grow
tall, while other species emphasise vertical
extension \citep{poorter-2006}.

The second term, ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$,
accounts for the marginal cost of deploying an additional unit of leaf
area, including construction of the leaf itself and various support
structures. As such, ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$
can itself be expressed as a sum of construction costs per unit leaf
area:

\begin{equation}\label{eq:daldmt}
\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}
= \big(\phi
 + \frac{{\rm d}M_\textrm{s}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}\big)^{-1}.
\end{equation}

The third term in eq \ref{eq:dhdt},
${\rm d}M_{\rm a} / {\rm d}P$, gives the fraction of net
biomass production (eq. \ref{eq:dbdt}) that is allocated to growth
rather than reproduction or storage.

In a similar way, basal area ($A_{\rm st}$) increment can be
expressed as the sum of increments in sapwood, bark \& heartwood areas
($A_\textrm{ss}, A_{\rm sb}, A_{\rm sh}$ respectively):
$\frac{{\rm d}A_{\rm st}}{{\rm d}t}= \frac{{\rm d}A_{\rm sb}}{{\rm d}t} + \frac{{\rm d}A_\textrm{ss}}{{\rm d}t} + \frac{{\rm d}A_{\rm sh}}{{\rm d}t}$.
We now have an equation for basal area growth that conatins many of the same
elements as eq. \ref{eq:dhdt}:

\begin{equation}\label{eq:dast}
\frac{{\rm d}A_{\rm st}}{{\rm d}t}=
\left(\frac{{\rm d}A_\textrm{ss}}{{\rm d}A_{\rm l}} + \frac{{\rm d}A_{\rm sb}}{{\rm d}A_{\rm l}}\right) \times
\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}} \times \frac{{\rm d}M_{\rm a}}{{\rm d}B} \times \frac{{\rm d}B}{{\rm d}t} +\frac{{\rm d}A_{\rm sh}}{{\rm d}t} .
\end{equation}

Diameter growth is then given by the geometric relationship between stem
diameter ($D$) and $A_{\rm st}$:

\begin{equation} \label{eq:dDdt}
\frac{{\rm d}D}{{\rm d}t}= \sqrt{\frac{\pi}{A_{\rm st}}} \frac{{\rm d}A_{\rm st}}{{\rm d}t}.
\end{equation}

@@

By definition, the main effect of increasing $\phi$ is to decrease the
marginal leaf deployment per mass (eq. \ref{eq:daldmt}) and thus reduce
growth rate. However, to fully account for the influence of $\phi$ on
growth we must also account for the decrease in leaf turnover that
results from superior leaf construction \citep{wright-2004}.
This is achieved by linking the turnover rate of leaf in eq.
\ref{eq:dbdt} to $\phi$. We use an allometric scaling relationship of
the form $k_{\rm l}=\alpha_4 \, \phi^{-\beta_4}$, which has been
shown to describe patterns across diverse
ecosystems \citep{wright-2004} (Fig. \ref{fS-leaf}).

@@

Unlike leaf and stem traits, height at maturation ($h_m$) moderates
growth by adjusting the amount of energy invested in growth, i.e.~the
term $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ in eqs.
\ref{eq:dhdt} and \ref{eq:dast}.

@@

Unlike eq. \ref{eq:RGR}, eqs. \ref{eq:dhdt}, \ref{eq:dast} and
\ref{eq:dDdt} predict a relationship between $\phi$ and growth that
changes with plant size ( \ref{f-lmA_growth_size}).
Decreasing $\phi$ has two impacts on growth rate. First, lower
$\phi$ increases marginal leaf deployment
(${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$ by economising on
construction costs. Second, lower $\phi$ decreases net production
(${\rm d}P / {\rm d}t$), due to increased leaf turnover. Whether
lower $\phi$ increases growth thus depends on the relative magnitude
of these two effects. When plants are small the effect on leaf
deployment rate is bigger and so decreasing $\phi$ increases growth
rate. When plants are large, the influence of $\phi$ on leaf
deployment rate is diminished, because the costs of building other
supportive tissues (other terms in eq. \ref{eq:daldmt}) are larger (Fig.
\ref{f-mass_fraction}). The net result is that at larger sizes, low
$\phi$ is no longer advantageous for growth (Fig.
\ref{f-lmA_growth_size}d).

@@

 Greater height at maturation ($h_m$)
can thus lead to a growth advantage by increasing
$\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ . At smaller sizes, there is no differentiation
among species, because most species are focusing on growth. At larger sizes, individuals of some species are
allocating a majority of their surplus energy to growth, leaving a
strong signal in potential growth rate.

@@

Current theory emphasises the relationship between $\phi$ and the relative growth in plant mass
\citep{lambers-1992, cornelissen-1996, wright_cross-2000,  enquist-2007},
${\rm d}P / {\rm d}t / M_{\rm a}$. This relationship can be derived from eq. \ref{eq:dbdt}  by ...
(see Appendix \ref{app:traits-RGR}).

@@

\begin{table}[ht]
 \caption{Key variables
%  $a = l + ss + sb +  r$,
%  $M_{sh}$,
%  $M_t = M_a + M_{sh}$,
%  $st = stem total$
}
\centering

@@

  $b$   & kg  & biomass originating from parent plant\\
  $M_i$ & kg  & mass retained on plant in tissue $i$\\
  $A_i$ & m$^2$  & surface-area or cross-sectional-area of tissue $i$\\
  $h$   & m  & height of plant\\
  $D$   & m  & diameter of plant stem\\

@@

{\footnotesize  % smaller font, double space
 \begin{doublespace}
  \begin{tabular}{p{2.5cm}p{3.5cm}p{5cm}p{4cm} }\\ \hline
  Variable & Function & Marginal cost & Growth rate\\ \hline
  \multicolumn{4}{l}{\textbf{a) Assumed relationship to leaf area}} \\ \\
  height &
    $h = \alpha_1 \, A_{\rm l}^{\beta_1}$ &
    $\frac{{\rm d}H}{{\rm d}A_{\rm l}} =  \beta_1 \alpha_1 A_{\rm l}^{\beta_1-1}$ &
    $\frac{{\rm d}H}{{\rm d}t}  = \frac{{\rm d}H}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
  sapwoood area &
    $A_\textrm{ss} = \theta \, A_{\rm l}$ &
    $\frac{{\rm d}A_\textrm{ss}}{{\rm d} A_{\rm l}} = \theta$ &
    $\frac{{\rm d}A_\textrm{ss}}{{\rm d}t}  =\frac{{\rm d}A_\textrm{ss}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
  bark area &
    $A_{\rm sb} = b \, \theta \, A_{\rm l}$ &
    $\frac{{\rm d}A_{\rm sb}}{{\rm d} A_{\rm l}} = b \, \theta$ &
    $\frac{{\rm d}A_{\rm sb}}{{\rm d}t} = \frac{{\rm d}A_{\rm sb}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\  \\
  \multicolumn{4}{l}{\textbf{b) Derived equation for mass of tissue }} \\
  leaf mass &
    $M_{\rm l} = \phi \, A_{\rm l} $ &
    $\frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}} = \phi$ &
    $\frac{{\rm d}M_{\rm l}}{{\rm d}t}  = \frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
  sapwood mass &
    $M_\textrm{ss} = \theta \, \rho \, \eta_c \, A_{\rm l} \, h $ &
    $\frac{{\rm d}M_\textrm{ss}}{{\rm d}A_{\rm l}} = \theta\, \rho\, \eta_c\, \big( h + A_{\rm l}\, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
    $\frac{{\rm d}M_\textrm{ss}}{{\rm d}t}  = \frac{{\rm d}M_\textrm{ss}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
  bark mass &
    $M_{\rm sb} = b\, \theta \, \rho \, \eta_c \, A_{\rm l} \, h $ &
    $\frac{{\rm d}M_{\rm sb}}{{\rm d}A_{\rm l}} = b \, \theta \, \rho \, \eta_c\big( h + A_{\rm l} \, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
    $\frac{{\rm d}M_{\rm sb}}{{\rm d}t}  = \frac{{\rm d}M_{\rm sb}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
  root mass &
    $M_{\rm r} = \alpha_3 \, A_{\rm l}$ &
    $\frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}} = \alpha_3$  &
    $\frac{{\rm d}M_{\rm r}}{{\rm d}t}  = \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\ \\

  \multicolumn{4}{l}{\textbf{c) Integrated quantities}} \\
  heartwood area &
    $A_{\rm sh} = \int_0^t \frac{{\rm d}A_{\rm sh}}{{\rm d}t}(t^\prime) \, dt^\prime$ &
     &
    $\frac{{\rm d}A_{\rm sh}}{{\rm d}t} = k_\textrm{s} \, A_\textrm{ss}$ \\
  heartwood mass &
    $M_{\rm sh} = \int_0^t \frac{{\rm d}M_{\rm sh}}{{\rm d}t}(t^\prime) \, dt^\prime$ &
     &
    $\frac{{\rm d}M_{\rm sh}}{{\rm d}t} = k_\textrm{s} \, M_\textrm{ss}$ \\
  \hline \\
\end{tabular}
NB -  theta is sapwood area per leaf area. In tree 2 has been opposite, but that makes notation more clumsy.
\end{doublespace}}
\label{tab:allometry}
\end{table}

@@

\begin{table}[ht]
\caption{Effect of traits and related trade-offs on demography}
\centering
{\footnotesize  % smaller font, double space
  \begin{doublespace}
  \begin{tabular}[c]{l|c|ccccc|c|c}
  \multicolumn{9}{c}{}\\ \hline
  & \multicolumn{8}{c}{\textbf{Effect of trait on}}\\
  & \textbf{Birth} & \multicolumn{5}{c|}{\textbf{Growth}} & \textbf{Mortality} & \textbf{Fecundity} \\
  & $h_0$
  & $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$
  & $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$
  & $\frac{{\rm d}B}{{\rm d}t}$
  & $\frac{{\rm d}A_\textrm{ss}}{{\rm d}A_{\rm l}}$
  & $\frac{1}{k_s}$ & &  \\\hline
  \textbf{This study}&&&&&&&&\\
  Leaf-construction cost, $\phi$ & & $\downarrow$ & $\uparrow$ & & & & \\
  Wood-construction cost, $\rho$ & & $\downarrow$ &  & & & $\downarrow$ & $\downarrow$  \\
  Height at maturation, $h_m$ & & &$\uparrow$ & & & & & $\downarrow$  \\
  &&&&&&&&\\\hline
  \textbf{Possible extensions}&&&&&&&&\\
  Sapwood area per leaf area, $\theta$ & & $\downarrow$& & $\uparrow$ & $\uparrow$ & & \\
  Leaf-nitrogen per area & & &$\downarrow$$\uparrow$ & & & & & \\
  Seed size & $\uparrow$ & & & & & & & $\downarrow$ \\ \hline
  \end{tabular}
  \end{doublespace}
}
\label{tab:trade-offs}
\end{table}

@@

We start with eq. \ref{eq:dbdt} for mass-based growth. For seedlings, which are
young and mostly leaf, it is reasonable to ignore all turnover terms as
well as the respiration terms for non-leaf tissues. Net production then
becomes a linear function of leaf area and net photosynthesis per leaf
area ($A_\textrm{net} = y(A - r_{\rm l})$), making relative growth
rate a linear function of $\phi$:

\begin{equation}\label{eq:RGR}
\underbrace{\strut\frac{{\rm d}B}{{\rm d}t}\frac{1}{M_{\rm a}}}_\textrm{relative growth in mass}  \approx A_\textrm{net} \times \phi^{-1} \times \frac{M_{\rm l}}{M_{\rm a}}. \end{equation}

Although Eq. \ref{eq:RGR} captures patterns of growth in seedlings in
relation to $\phi$\citep{wright_cross-2000}, this
approximation does not directly link to other traits, or to the
variables that are routinely collected for large trees: namely plant
height ($h$) and stem cross-sectional area ($A_{\rm st}$) or
diameter $D$.

@@

\begin{equation}\label{eq:ha}
A_{\rm l}=\alpha_1 \, h^{\beta_1}.
\end{equation}

@@

\begin{equation}\label{eq:crown2}
\frac{M_{\rm l}}{\phi}= \theta \, S(0,h).
\end{equation}

@@

\begin{equation}\label{eq:ms1}
M_\textrm{s}=\rho \, \int_0^h \, S(z,h) \, {\rm d}z= \rho \, S(0,h) \, h \, \eta_c, \end{equation}

@@

\begin{equation}\label{eq:ms2}
M_\textrm{s}=\rho \, \eta_c \, \theta \, A_{\rm l} \, h.
\end{equation}

@@

\begin{equation}\label{eq:mb}
M_{\rm b}=b M_\textrm{s}.
\end{equation}

@@

\begin{equation}\label{eq:mr}
M_{\rm r}=\alpha_3 \, A_{\rm l}.
\end{equation}
